---
title: "Summary table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summary table}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = TRUE,
  warning=FALSE,
  fig.height = 6,
  fig.width = 9,
  fig.align='center'
)
```

```{r color-function, echo = FALSE}
colorize <- function(text, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, text)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, text)
  } else text
}

```

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyverse)
library(SAKK)
library(gtsummary)

source("../R/summaryTable.R")
```

The function `summaryTable()` produces a table with descriptive
statistics for continuous, categorical and dichotomous variables. It is
largely based on the function `gtsummary::tbl_summary()`. The changes as
compared to `tbl_summary` are:

-   Simplification of syntax
-   Missing values for categorical variables can be counted in the
    percentage.

## Setup and data

To demonstrate the various functionalities of the function we will use
the dataset `survival::colon`.

```{r, message = FALSE}
library(survival)
data(cancer, package="survival")

colon1 <-  colon %>%
  group_by(id) %>%
  slice(1) %>% # Select the first row within each id group
  ungroup()
  
```

```{r, echo = FALSE}
n_patients <- nrow(colon)
```

The dataset `colon` contains data of `r n_patients` patients from one of
the first successful trials of adjuvant chemotherapy for colon cancer.

For simplicity, we focus here on recurrence only, two treatment groups,
and four variable: the treatment group (`rx`), the sex (`sex`), the age
(`age` and `sex_char`) and the extent of local spread (`extent`). We
also add a few missing values for the variable `extent`.

```{r}
set.seed(123)
colon2 <- colon1 %>%
  select(rx, sex, age, extent) %>%
  filter(rx != "Lev") %>%
  mutate(sex_char = if_else(sex == 1, "male", "female"),
         extent = if_else(row_number() %in% sample(row_number(), size = round(0.1 * n())), NA, extent))

colon2$rx <- droplevels(colon2$rx)

```

```{r}
head(colon2)
```

## Simple table

By default, the function produces a table with all variables present in
the dataset.

```{r}
summaryTable(data = colon2)
```

**Currently, the name of the different variables cannot be modified in
the function directly, they have to be modified in the dataset. OK?**

```{r}
colon2 <- colon2 %>% 
  rename(Arm = rx, Sex = sex, Age = age, Extent = extent) %>% 
  mutate(Extent = as.factor(Extent))
  
```

```{r}
summaryTable(colon2)
```

## Stratify per group

The argument `group` indicates by which variable the summary statistics
will be stratified.

```{r}
summaryTable(data = colon2, group = "Arm")
```

## Continuous variables

If only specific variables are to be included in the table, they can be
written in the argument `vars`.

```{r}
summaryTable(data = colon2, group = "Arm", vars = "Age")
```

By default, the function plots the median and (Q1, Q3) for continuous
variables. A number of other options are available, using the argument
`stat_cont`.

### Median and range

```{r}
summaryTable(data = colon2, group = "Arm", vars = "Age", 
             stat_cont = "median_range")
```

### Mean and standard deviation

```{r}
summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Age", 
             stat_cont = "mean_sd")
```

## Tests

By default, no p-value and confidence (CI) are displayed. p-values can
be added with `test_cont` and CI with `ci_cont`.

```{r}
summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Age", 
             stat_cont = "mean_sd", 
             test_cont = "t.test",
             ci_cont = "t.test"
             )
```

## Categorical variables

For categorical variables, the percentage are not automatically added
next to the missing. They can be added by setting the argument `missing`
to `TRUE`.

```{r}
summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Extent", 
             test_cat = "chisq.test", 
             missing = TRUE
             )


```

Note that the p-value is always calculated ignoring the missing values.

Both can also be put next to each other:

```{r}
summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Extent", 
             test_cat = "chisq.test", 
             missing = "both"
             )

```

## Binary variables

By default, both levels are showed for binary variables. Setting the
argument `binary` to `TRUE` only shows one level:

```{r}
summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Sex"
             )

summaryTable(data = colon2, 
             group = "Arm", 
             vars = "Sex", 
             binary = TRUE)

```

## Further customization 

Digits can be customized with the arguments `digits_cont`and
`digits_cat`.

# To do: 

-   Adding a column "total"

-   summaryByVisit and summaryLevel?

-   Time-to-event endpoints?
