---
title: "Summary table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summary table}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = TRUE,
  warning=FALSE,
  fig.height = 6,
  fig.width = 9,
  fig.align='center'
)
```

```{r color-function, echo = FALSE}
colorize <- function(text, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, text)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, text)
  } else text
}

```

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyverse)
library(gtsummary)
library(summarySCI)
library(flextable)
```

The function `summaryTable()` produces a table with descriptive
statistics for continuous, categorical and dichotomous variables. It is
largely based on the function `gtsummary::tbl_summary()`. The changes as
compared to `tbl_summary` are:

-   Simplification of syntax
-   Missing values for categorical variables can be counted in the
    percentage.
- Possibility to add the number of observation per group as a column.

## Setup and data

To demonstrate the various functionalities of the function we will use
the dataset `survival::colon`.

```{r, message = FALSE}
library(survival)
data(cancer, package="survival")

colon1 <-  colon %>%
  group_by(id) %>%
  slice(1) %>% # Select the first row within each id group
  ungroup()
  
```

```{r, echo = FALSE}
n_patients <- nrow(colon)
```

The dataset `colon` contains data of `r n_patients` patients from one of
the first successful trials of adjuvant chemotherapy for colon cancer.

For simplicity, we focus here on recurrence only, two treatment groups,
and four variable: the treatment group (`rx`), the sex (`Male`), the age
(`age`) and the extent of local spread (`extent`). We also add a few
missing values for the variable `extent`.

```{r}
set.seed(123)
colon2 <- colon1 %>%
  select(rx, sex, age, extent) %>%
  filter(rx != "Lev") %>%
  mutate(rx = if_else(rx == "Obs", "Control", rx),
         extent = if_else(row_number() %in% sample(row_number(), size = round(0.1 * n())), NA, extent)) %>% 
  rename(Male = sex) %>% 
  mutate(extent = as.factor(extent))


```

```{r}
head(colon2)
```

## Simple table

By default, the function produces a table with all variables present in
the dataset.

```{r}
summaryTable(data = colon2)
```
If only specific variables are to be included, the need to be entered 
in the argument `vars`. The argument `group` allows the summary 
statistics to be stratified by this variable. 

```{r}
summaryTable(data = colon2, 
             vars = c("Male", "age", "extent"), 
             group = "rx")
```


### Displayed name of variables

The displayed name of each variable is

-   the label if it exists in the dataset, or

-   the variable name if no label is present in the dataset (which is
    the case in our example).

In order to customize the displayed name, the argument `labels` can be
used. Please note that it needs to be entered as a list, as shown below:

```{r}
summaryTable(data = colon2, 
                 labels = list(rx = "Arm", age = "Age", extent = "Extent"))
```


## Adding number of observations 

The number of observations which are not missing values can be added as a column setting the 
argument `add_n` to `TRUE`.

```{r}
summaryTable(data = colon2, 
                 labels = list(rx = "Arm", age = "Age", extent = "Extent"), 
             add_n = TRUE)
```




## Stratify per group

The argument `group` indicates by which variable the summary statistics
will be stratified.

```{r}
summaryTable(data = colon2, group = "rx", 
             labels = list(age = "Age", extent = "Extent"))
```

An "overall" column can be added by setting the argument `overall` to
`TRUE`.

```{r}
summaryTable(data = colon2, group = "rx", overall = TRUE, 
              labels = list(age = "Age", extent = "Extent"))
```

## Variable types

Per default, numeric variables are considered continuous, unless 
they only have two unique values: 0 and 1. In that case, they are considered 
dichotomous.
Continuous variable can be displayed as continuous with 
`continuous_as = continuous` (default) but can also be displayed as categorical 
with `continuous_as = categorical`. 
For dichotomous variables, all levels are displayed per default 
(`dichotomous_as = categorical`), but only one row can be shown using `dichotomous_as = dichotomous`. In that case, the reference value is chosen with the argument 
`value = list(variable ~ "level to show")`. 



```{r}
summaryTable(data = colon2, group = "rx", vars = "age",
              labels = list(age = "Age"))
```

By default, the function plots the median and range for continuous
variables. A number of other options are available, using the argument
`stat_cont`.

### Stat

The statistics to be displayed can be chosen using `stat_cont` and
`stat_cat`, for example: 

```{r}
summaryTable(data = colon2, group = "rx", vars = "age", 
             stat_cont = "median_IQR", 
              labels = list(age = "Age"))
```

```{r}
summaryTable(data = colon2, 
             group = "rx", 
             vars = "age", 
             stat_cont = "mean_sd", 
              labels = list(age = "Age"))
```

## Tests

By default, no p-value and confidence (CI) are displayed. p-values can
be added
by setting `test` to `TRUE` and CI by setting `ci` to `TRUE`.

The default test and CI types for continuous variable is `wilcox.test`.
This can
be changed in `test_cont` and `ci_cont`, respectively.

```{r}
summaryTable(data = colon2, 
             group = "rx", 
             vars = "age", 
             stat_cont = "mean_sd", 
             test = TRUE,
             ci = TRUE,
             labels = list(age = "Age")
             )
```


For categorical variables, the percentage are automatically added next
to the missing. This can be disabled by setting the argument `missing`
to `FALSE`.

```{r}
summaryTable(data = colon2, 
             group = "rx", 
             vars = "extent", 
             test = TRUE,
             ci = TRUE,
             missing_percent = TRUE, 
             labels = list(extent = "Extent")
             )


```

Note that the p-value is always calculated ignoring the missing values.\
For categorical variables, the default test type is `chisq.test` and the
default CI types is `wilson`. This can be changed in `test_cat` and
`ci_cat`, respectively.

The tables with and without missing values can also be put next to each
other
by setting `missing` to `"both"`.

```{r}
summaryTable(data = colon2, 
             group = "rx", 
             vars = "extent", 
             missing_percent = "both", 
              labels = list(extent = "Extent")
             )

```



## Further customization

Digits can be customized with the arguments `digits_cont`and
`digits_cat`.

# Further steps:

-   Adding time-to-event endpoints
-   New functions: summaryByVisit and summaryLevel

-   Adding tests
